---
title: "HW_Review"
author: "Yichien Chou"
date: "2020/3/11"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Problem 1
##Part A
Since there are two factors (alcoholic, treatment). Therefore, it is a Two-Way basic factorial design.

## Part B
It is a Split Plot design. The whole plot factor is coencentration of auxin; The whole plot is the plants; The split plot factor is whether the pair was debladed; The splt plot is the three lowerest leaf pairs.

## Part C
One-Way Basic Factorial Design.

## Part D
Since it splits experimental units into groups(block) of similar units, then assign treatment to units separately with each block. Thus, it is a Complete Block Design.

# Problem 2

```{r,echo=FALSE, message=FALSE}
library(ggplot2)
library(lsmeans)
options(contrasts =c("contr.sum", "contr.poly"))
```
## Part A

```{r, echo=FALSE}
cook <- read.csv('/Users/jason13nn/Desktop/SMU/Spring 2020/STAT 6302 (Experimental Statistics)/files/cook.csv')
head(cook)
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  library(plyr)
  
  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm) sum(!is.na(x))
    else       length(x)
  }
  
  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )
  
  # Rename the "mean" column    
  datac <- rename(datac, c("mean" = measurevar))
  
  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
  
  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult
  
  return(datac)
}

cook.summary <- summarySE(cook, measurevar="iron", groupvars=c("pot", "dish"))
cook.summary
```

## Part B
```{r, echo=FALSE}
ggplot(cook.summary, aes(x=pot, y=iron, group=dish, color=dish)) + 
  geom_errorbar(aes(ymin=iron-se, ymax=iron+se), width=0.1, 
                position=position_dodge(0.1)) +
  geom_line(position=position_dodge(0.1)) +
  geom_point(position=position_dodge(0.1)) +
  theme_bw()
```

## Part C
```{r, echo=FALSE}
cook.anova <- aov(iron ~ pot + dish + pot*dish, data=cook)
summary(cook.anova)
```

``````{r, echo=FALSE}
boxplot(iron ~ pot + dish, data=cook)
```

```{r, echo=FALSE}
par(mfrow=c(2,2))
plot(cook.anova)
```
We assume the observations are independent and samples are randomly selected. From QQ plot, it seems satisfied normality assumption. However, from boxplot, equal standard deviation assumption is violated. 
There is evidence of an interaction (F=4.89, p=0.00425), as well as both main effects (F=92.26, p<0.0001; F=34.46, p<0.0001). However, since the interaction is significant we will do post hoc comparisons on the crossedfactors.

# Part D
```{r, echo=FALSE}
lsmeans(cook.anova, pairwise ~ pot + dish)
```
In sum, aluminum and clay don’tappear to be different unless the vegetable version is compared to the other twofoods.

# Problem 3
```{r, echo=FALSE}
library(factoextra)
```
## Part A
```{r, echo=FALSE}
glass <- read.csv("/Users/jason13nn/Desktop/SMU/Spring 2020/STAT 6302 (Experimental Statistics)/files/glass.csv")

glassPCA <- prcomp(glass[,1:9], center=T, scale=T)
summary(glassPCA)
```

## Part B
```{r, echo=FALSE}
fviz_eig(glassPCA)
```

```{r, echo=FALSE}
(VE <- glassPCA$sdev^2)
PVE <- VE / sum(VE)
round(PVE, 2)
```

The first four PCs explained 0.28 + 0.23 + 0.16 + 0.13 = 80% of the variance.
Therefore, we keep the first four principal components. 

## Part C
```{r, echo=FALSE}
fviz_pca_var(glassPCA, col.var = "contrib", 
gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
repel = TRUE)
```
```{r, echo=FALSE}
glassPCA$rotation[,1:4]
```
PC1 shows that `RI`, `Al`, and `Ca` placed most weight on the first loading. It is hard to name this component due to uncertainty of their relationship. 

## Part D
```{r, echo=FALSE}
fviz_pca_ind(glassPCA, label="none", habillage=glass$Type,
addEllipses=TRUE, ellipse.level=0.95, palette = "Dark2")
```

It is not clear to see a separation of groups. Thus, we should also use PC3 and PC4 to classify glass type.
