---
title: "HW5"
author: "Yichien Chou"
date: "2020/3/29"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning= FALSE}
library(MASS)
library(caret)
library(klaR)
library(pROC)
library(OptimalCutpoints)

bird <- read.csv("/Users/jason13nn/Desktop/SMU/Spring 2020/STAT 6302 (Experimental Statistics)/files/Bird.csv")

levels(bird$LC)

#Define reference
bird$LC <- relevel(bird$LC, ref = "NoCancer")
```

## Part A


```{r, echo=FALSE, message=FALSE, warning= FALSE}
bird.lm <- glm(LC ~ FM + SS + BK + AG + YR + CD, family = "binomial", data = bird)
summary(bird.lm)
```

## Part B

*BK*: The objects with no bird multiplies the odds of no Lung Cancer by exp(-1.363) = 0.256 times.
*YR*: An one unit increases in YR multiplies the odds of no Lung Cancer by exp(0.723) = 0.485.

## Part C

```{r, echo=FALSE, message=FALSE, warning= FALSE}
##Predict the test set
bird$Cancer <- ifelse(bird$LC == "NoCancer",0,1) 
bird$pred.prob <- predict(bird.lm , bird, type = "response")
bird$Pred.Group <- ifelse(bird$pred.prob>0.5,"LungCancer","NoCancer") 

head(bird)
```

```{r, echo=FALSE, message=FALSE, warning= FALSE}
bird$Pred.Group <- as.factor(bird$Pred.Group)
bird$LC <- relevel(bird$LC, ref = "LungCancer")

##Create the confusion matrix from the test set.
confusionMatrix(data = bird$Pred.Group, 
                reference = bird$LC)
```

```{r, echo=FALSE, message=FALSE, warning= FALSE}
##ROC curves
birdROC <- roc(bird$LC, bird$pred.prob)

plot(birdROC, legacy.axes = TRUE)

summary(optimal.cutpoints(X="pred.prob", status="LC", data=bird, 
                          tag.healthy='NoCancer', methods='Youden'))
```

## Part D

```{r, message=FALSE, warning= FALSE}
#Brier Score
library(scoring)

brierScore <- mean((bird$pred.prob-bird$Cancer)^2)
brierScore
```

## Part E

Classification error is insensitive and statistically inefficient (In this case, the cutpoint is 0.5, then 0.501 and 0.999 are treated the same).

## Bonus

*(A)* It is opposite to the 45 degree line.
*(B)* To correct the issue, we should change the reference of `LC` to "LungCancer".
