---
title: "Final Project"
author: "Yichien Chou"
date: "2020/4/28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Description

Expedia provided logs of customer behavior. These include what customers searched for, how they interacted with search results (click/book), whether or not the search result was a travel package. More details can be found [here](https://www.kaggle.com/c/expedia-hotel-recommendations/data).

The data set has 37,670,293 observations and 24 variables. 

# Data Preparation

```{r, echo=FALSE}
#Use ff package to read the large data set

library(ff)
hotel <-read.csv.ffdf(file = "/Users/jason13nn/Desktop/SMU/Spring 2020/STAT 6302 (Experimental Statistics)/Final Project/expedia-hotel-recommendations/expedia_hotel.csv")
#Transform back to data frame
hotel <- as.data.frame(hotel)

dim(hotel)
#37670293 24

#randomly select 12,000 obs from the data set (6,000 for is_booking = 0, and 6,000 for is_booking = 1) for training set.
set.seed(123)
hotel_sampled1 <- hotel[sample(which(hotel$is_booking >0), 6000),]
hotel_sampled2 <- hotel[sample(which(hotel$is_booking <1), 6000),]
hotel.sample <- rbind(hotel_sampled1, hotel_sampled2)

#randomly select 10,000 obs to the training set
set.seed(999)
hotel.train <- hotel.sample[sample(1:nrow(hotel.sample), 10000),]

#The remained 2,000 obs to the test set
hotel.test <- setdiff(hotel.sample, hotel.train)
```

# Pre-processing

```{r, echo=FALSE}
#See the number of NAs for each variable
sapply(hotel.train, function(x) sum(is.na(x)))
#3,492 NAs in orig_destination_distance

#Impute missing values
library(VIM)
hotel.train$orig_destination_distance <- hotdeck(hotel.train$orig_destination_distance)
hotel.train$orig_destination_distance.data_imp <- NULL
hotel.train$orig_destination_distance <- hotel.train$orig_destination_distance$data

#Factorize predictor
hotel.train$site_name <- as.factor(hotel.train$site_name)
hotel.train$posa_continent <- as.factor(hotel.train$posa_continent)
hotel.train$user_location_country <- as.factor(hotel.train$user_location_country)
hotel.train$user_location_region <- as.factor(hotel.train$user_location_region )
hotel.train$user_location_city <- as.factor(hotel.train$user_location_city)
hotel.train$is_mobile <- as.factor(hotel.train$is_mobile)
hotel.train$is_package <- as.factor(hotel.train$is_package)
hotel.train$channel <- as.factor(hotel.train$channel)
hotel.train$srch_destination_id <- as.factor(hotel.train$srch_destination_id)
hotel.train$srch_destination_type_id <- as.factor(hotel.train$srch_destination_type_id)
hotel.train$is_booking <- as.factor(hotel.train$is_booking)
hotel.train$hotel_continent <- as.factor(hotel.train$hotel_continent)
hotel.train$hotel_country <- as.factor(hotel.train$hotel_country)
hotel.train$hotel_market <- as.factor(hotel.train$hotel_market)
hotel.train$hotel_cluster <- as.factor(hotel.train$hotel_cluster)

#date predictors (Keep srch_ci and transform into year and month)
hotel.train$srch_ci <- as.Date(hotel.train$srch_ci)
hotel.train$srch_ci <- as.POSIXlt(hotel.train$srch_ci)
hotel.train$year <- hotel.train$srch_ci$year + 1900
hotel.train$month  <- hotel.train$srch_ci$mon  + 1

hotel.train <- na.omit(hotel.train)
#9,994 obs left

#There were 15 categorical variables (including response), 3 date variables, and 5 continuous variables, the 24th was ID.

#Repeat on test set
sapply(hotel.test, function(x) sum(is.na(x)))
#730 NAs

hotel.test$orig_destination_distance <- hotdeck(hotel.test$orig_destination_distance)
hotel.test$orig_destination_distance.data_imp <- NULL
hotel.test$orig_destination_distance <- hotel.test$orig_destination_distance$data

#Factorize predictor
hotel.test$site_name <- as.factor(hotel.test$site_name)
hotel.test$posa_continent <- as.factor(hotel.test$posa_continent)
hotel.test$user_location_country <- as.factor(hotel.test$user_location_country)
hotel.test$user_location_region <- as.factor(hotel.test$user_location_region )
hotel.test$user_location_city <- as.factor(hotel.test$user_location_city)
hotel.test$is_mobile <- as.factor(hotel.test$is_mobile)
hotel.test$is_package <- as.factor(hotel.test$is_package)
hotel.test$channel <- as.factor(hotel.test$channel)
hotel.test$srch_destination_id <- as.factor(hotel.test$srch_destination_id)
hotel.test$is_booking <- as.factor(hotel.test$is_booking)
hotel.test$srch_destination_type_id <- as.factor(hotel.test$srch_destination_type_id)
hotel.test$hotel_continent <- as.factor(hotel.test$hotel_continent)
hotel.test$hotel_country <- as.factor(hotel.test$hotel_country)
hotel.test$hotel_market <- as.factor(hotel.test$hotel_market)
hotel.test$hotel_cluster <- as.factor(hotel.test$hotel_cluster)

#date predictors (Keep srch_ci and transform into year and month)
hotel.test$srch_ci <- as.Date(hotel.test$srch_ci)
hotel.test$srch_ci <- as.POSIXlt(hotel.test$srch_ci)
hotel.test$year <- hotel.test$srch_ci$year + 1900
hotel.test$month  <- hotel.test$srch_ci$mon  + 1

hotel.test <- na.omit(hotel.test)
#1,999 obs left
```

# EDA

```{r, echo=FALSE}
#EDA
library(ggplot2)
#1. Site_name (No significant difference)
#ggplot(hotel.train) +
#  geom_bar(aes(x = site_name, fill = is_booking))

#2. posa_continent (No significant difference)
# ggplot(hotel.train) +
# geom_bar(aes(x = posa_continent, fill = is_booking)) +
#  ggtitle("Barplot of posa_continent with booking") +
#  xlab("Continent associated with the Expedia point of sale") +
#  scale_x_discrete(labels = c("Continent 0", "Continent 1", "Continent 2", "Continent 3", "Continent 4")) +
#  scale_fill_discrete(name = "Booking sucessfully?", labels = c("Yes", "No"))

#3. is_mobile (No significant difference)
# ggplot(hotel.train) +
#    geom_bar(aes(x = is_mobile, fill = is_booking)) +
#    ggtitle("Barplot of is_mobile with booking") +
#    xlab("Is users connect from a mobile device?") +
#    scale_fill_discrete(name = "Booking sucessfully?", labels = c("Yes", "No"))

#4. is_package
ggplot(hotel.train) +
    geom_bar(aes(x = is_package, fill = is_booking)) +
    ggtitle("Barplot of is_package with booking") +
    xlab("Is the booking was generated as a package (combined with a flight)?") +
    scale_fill_discrete(name = "Booking sucessfully?", labels = c("Yes", "No"))

##It seems it is more like to book the hotel if the booking was generated as a part of a package.

#5. channel (No significant difference)
#ggplot(hotel.train) +
#    geom_bar(aes(x = channel, fill = is_booking)) +
#    ggtitle("Barplot of channel with booking") +
#    xlab("ID of a marketing channel") +
#    scale_fill_discrete(name = "Booking sucessfully?", labels = c("Yes", "No"))

#6. srch_destination_type_id
ggplot(hotel.train) +
    geom_bar(aes(x = srch_destination_type_id, fill = is_booking)) +
    ggtitle("Barplot of srch_destination_type_id with booking") +
    xlab("Type of destination") +
    scale_fill_discrete(name = "Booking sucessfully?", labels = c("Yes", "No"))

##The proportion of booking seems different between each type of destination.

#7. hotel_continent
ggplot(hotel.train) +
    geom_bar(aes(x = hotel_continent, fill = is_booking)) +
    ggtitle("Barplot of hotel_continent with booking") +
    xlab("Hotel continent") +
    scale_fill_discrete(name = "Booking sucessfully?", labels = c("Yes", "No"))

##The probability of booking in hotel continent 2 seems less than the others.

#8. hotel_cluster
ggplot(hotel.train) +
    geom_bar(aes(x = hotel_cluster, fill = is_booking)) +
    ggtitle("Barplot of hotel_cluster with booking") +
    xlab("ID of a hotel cluster") +
    scale_fill_discrete(name = "Booking sucessfully?", labels = c("Yes", "No"))

##It seems the probability of booking was differ from hotel clusters.

#9.  orig_destination_distance (No significant difference)
#ggplot(hotel.train, aes(x=is_booking, y=orig_destination_distance,color=is_booking)) + 
#    geom_boxplot()+
#    scale_color_manual(values=c("#999999", "#56B4E9"))+
#    labs(title="Boxplot of physical distance with booking", x="Booking", y = "Physical distance") +
#    scale_x_discrete(labels = c("Yes", "No"))

#10. srch_adults_cnt
ggplot(hotel.train) +
    geom_bar(aes(x = as.factor(srch_adults_cnt), fill = is_booking)) +
    ggtitle("Barplot of number of adults in the hotel room") +
    xlab("Number of adults in the hotel room") +
    scale_fill_discrete(name = "Booking sucessfully?", labels = c("Yes", "No"))

##If only one adult in the room, the probability of booking was lower than others.

#11. srch_children_cnt (No significant difference)
#ggplot(hotel.train) +
#    geom_bar(aes(x = as.factor(srch_children_cnt), fill = is_booking)) +
#    ggtitle("Barplot of number of children in the hotel room") +
#    xlab("Number of children in the hotel room") +
#    scale_fill_discrete(name = "Booking sucessfully?", labels = c("Yes", "No"))

#12. srch_rm_cnt (No significant difference)
#ggplot(hotel.train) +
#    geom_bar(aes(x = as.factor(srch_rm_cnt), fill = is_booking)) +
#    ggtitle("Barplot of number of rooms in the search") +
#    xlab("Number of rooms in the search") +
#    scale_fill_discrete(name = "Booking sucessfully?", labels = c("Yes", "No"))

#13. cnt
ggplot(hotel.train) +
    geom_bar(aes(x = as.factor(cnt), fill = is_booking)) +
    ggtitle("Barplot of the number of similar events in the context of the same user session") +
    xlab("Number of similar events in the context of the same user session") +
    scale_fill_discrete(name = "Booking sucessfully?", labels = c("Yes", "No"))

#The probability of booking seems differnt in cnt, but it is a umbalanced variable.

#14. year
ggplot(hotel.train) +
  geom_bar(aes(x = as.factor(year), fill = is_booking)) +
  scale_fill_brewer(palette = "Set1")

#The year might also be one of factors.

#15. month (No significant difference)
#ggplot(hotel.train) +
#  geom_bar(aes(x = as.factor(month), fill = is_booking)) +
#  scale_fill_brewer(palette = "Set1")
```

# Logistic Regression

```{r, echo=FALSE, warning=FALSE}
##Logistic
#Remove predictors
hotel.train$date_time <- NULL
hotel.train$user_id <- NULL
hotel.train$srch_ci <- NULL
hotel.train$srch_co <- NULL

hotel.test$date_time <- NULL
hotel.test$user_id <- NULL
hotel.test$srch_ci <- NULL
hotel.test$srch_co <- NULL

#build logistic model
hotel.train.logit <- glm(is_booking ~ ., data = hotel.train[,-c(2:5,13,18:20)], family='binomial')
summary(hotel.train.logit)

#find optimal cutoff
prob <- predict(hotel.train.logit, hotel.test, type="response")  # predicted scores

library(InformationValue)
optCutOff <- optimalCutoff(actuals = hotel.test$is_booking, predictedScores =  prob)[1]
optCutOff
#The optimal cutoff is 0.4752473.

#confusion matrix
confusionMatrix(hotel.test$is_booking, prob, threshold = optCutOff)

#Misclass error
misClassError(actuals = hotel.test$is_booking, predictedScores = prob, threshold = optCutOff)
#0.3287

#Test Accuracy
precision(hotel.test$is_booking, prob, threshold = optCutOff)
#0.6173657

#ROC
ROC <- plotROC(actuals = as.factor(hotel.test$is_booking), predictedScores =  prob)
ROC

#Residuals plot
Index <- 1:dim(hotel.train)[1]
# deviance residuals
Deviance_Residuals <- residuals(hotel.train.logit)
dff <- data.frame(Index,Deviance_Residuals,hotel.train$is_booking)
 
ggplot(data = dff, mapping = aes(x = Index,y = Deviance_Residuals,color = hotel.train$is_booking)) +
  geom_point() +
  geom_hline(yintercept = 3,linetype = "dashed", color = "blue") +
  geom_hline(yintercept = -3,linetype = "dashed", color = "blue") +
  labs(title = "Plot of Deviance Residuals") +
  theme(plot.title = element_text(hjust = 0.5))
#Most of the residuals falls with (-3, 3).
```

# Machine Learning Methods

```{r, echo=FALSE}
#Neural Network
library(parallel) 
library(doParallel)
start_time2 <- Sys.time()
cl <- makePSOCKcluster(3) 
registerDoParallel(cl)

library(nnet)
hotel.nn <- nnet(is_booking ~ ., data = hotel.train[,-c(2:5,13,18:20)], size = 10)
nn.predict <- predict(object = hotel.nn, newdata = hotel.test, type = "raw")

table.nn <- table(nn.predict, hotel.test$is_booking)
table.nn

#Test Accuracy
(table.tree[1,1]+table.tree[2,2])/sum(table.tree)

#CART
library(rpart)
hotel.tree <- rpart(is_booking ~ ., data = hotel.train[,-c(2:5,13,18:20)], method = "class", cp = 1e-3)

tree.predict <- predict(hotel.tree, hotel.test, type = "class")

table.tree <- table(tree.predict, hotel.test$is_booking)
table.tree

#Test Accuracy
(table.tree[1,1]+table.tree[2,2])/sum(table.tree)
#0.6573287

#Random Forest
library(randomForest)
hotel.rf <- randomForest(is_booking ~ . , hotel.train[,-c(2:5,13,18:20)], n_tree = 100)

rf.predict <- predict(hotel.rf, hotel.test, type = "class")

table.rf <- table(rf.predict, hotel.test$is_booking)
table.rf

#Test Accuracy
(table.tree[1,1]+table.tree[2,2])/sum(table.tree)
#

#SVM


stopCluster(cl) 
end_time2 <- Sys.time() 
end_time2 - start_time2


